Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections at the LArIAT experiment

### 2018-06-01 Chapter-1/1-Theory.tex
s

A summary of the pion final states in order of pion multiplicity for the reaction channel is given in table \ref{tab:pionFinal}. Pion capture and pion decay at rest dominate the cross section under 100 MeV. We define pion capture as the process determining the formation of a pionic atom and the subsequent pion's end of life. Stopping negative pions can form pionic argon, where the negative pion plays the role of an orbital electron. Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave form of the pion will overlap more with the nucleus compared to the electron case. After the electromagnetic formation of the pionic atom, the pion will get quickly absorbed by the nucleus, which is put in an excited state. The nucleus then de-excites with the emission of low energy nucleons and photons.  Pion capture is dominant compared to pion decay, the other important process for very low energy pions. The decay of a pion is governed by the weak force; the pion decay life time is $\tau_\pi = 2.6 \times 10^{-8}$ s and the main decay mode is  $\pi^-\rightarrow \mu^- + \bar\nu_\mu$ (BR 99.98\%). %; for pion decay at rest, the muon is emitted with an energy of $\sim5$ MeV and it decays into $\mu^-\rightarrow e^- + \bar\nu_e + \nu_\mu$ in $\tau_\mu = 2.3 $ $\mu$s. 
 Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.

@@

>>>>

A summary of the pion final states in order of pion multiplicity for the reaction channel is given in table \ref{tab:pionFinal}. Figure \ref{fig:PionsEvd} shows examples of the topologies due to the pion-argon hadronic interaction as they appear in the  LArIAT data: elastic and inelastic scattering, pion absorption with emission of protons, charge exchange and production of pions.

\begin{figure}[hbpt]
\centering
\includegraphics[width=\textwidth]{Chapter-1/Images/PionsEvd.pdf}
\caption{Pion-Argon interaction topologies due to elastic scattering, inelastic scattering, pion absorption with emission of protons, charge exchange and production of pions as seen in LArIAT data (induction plane only).}
\label{fig:PionsEvd}
\end{figure}

Pion capture and pion decay at rest dominate the cross section under 100 MeV. We define pion capture as the process determining the formation of a pionic atom and the subsequent pion's end of life. Stopping negative pions can form pionic argon, where the negative pion plays the role of an orbital electron. Since the pion mass is two orders of magnitude greater than the electron mass, the spatial wave form of the pion will overlap more with the nucleus compared to the electron case. After the electromagnetic formation of the pionic atom, the pion will get quickly absorbed by the nucleus, which is put in an excited state. The nucleus then de-excites with the emission of low energy nucleons and photons.  Pion capture is dominant compared to pion decay, the other important process for very low energy pions. The decay of a pion is governed by the weak force; the pion decay life time is $\tau_\pi = 2.6 \times 10^{-8}$ s and the main decay mode is  $\pi^-\rightarrow \mu^- + \bar\nu_\mu$ (BR 99.98\%). %; for pion decay at rest, the muon is emitted with an energy of $\sim5$ MeV and it decays into $\mu^-\rightarrow e^- + \bar\nu_e + \nu_\mu$ in $\tau_\mu = 2.3 $ $\mu$s. 
 Since pion capture can be considered an electromagnetic process and pion decay is a weak process, this energy region is purposely excluded from the hadronic cross section measurement.

@@

Figure \ref{fig:KaonsEvd} shows examples of the topologies due to the kaon-argon hadronic interaction as they appear in the  LArIAT data: elastic and inelastic scattering.

\begin{figure}[hbpt]
\centering
\includegraphics[width=0.7\textwidth]{Chapter-1/Images/KaonsEvd.pdf}
\caption{Kaon-Argon interaction topologies due to elastic scattering and inelastic scattering as seen in LArIAT data (collection plane only).}
\label{fig:KaonsEvd}
\end{figure}

### 2018-06-01 Chapter-7/7-KaonCrossSection.tex
s

The measured cross section is mostly in disagreement with the predicted cross section; in particular it is significantly higher -- on the order of 20\%-30\% higher -- in the bins lower than 400 MeV. Such difference is unlikely to be accounted for the small proton contamination expected in the data sample. It is worth noticing that the 100 MeV to 400 MeV energy range is the region of interest for the expected kaons coming from a proton decay event. With this result, we will inform a tuning of the Geant4 model for kaon cross section.

>>>>

The measured cross section is mostly in tension with the predicted cross section; in particular it is significantly higher -- on the order of 20\%-30\% higher -- in the bins lower than 400 MeV. Such difference is unlikely to be accounted for the small proton contamination expected in the data sample. It is worth noticing that the 100 MeV to 400 MeV energy range is the region of interest for the expected kaons coming from a proton decay event. With this result, we will inform a tuning of the Geant4 model for kaon cross section.

### 2018-06-01 thesis_main.tex
s

This thesis presents two world's first measurements: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1050 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-650 MeV kinetic energy range. The analyses devised for these measurements use both the core elements of LArIAT: beamline and TPC. The first step in each analysis is the development of an event selection based on  beamline and TPC information geared towards the identification of the hadron of interest. We then proceed to match the beamline candidate to a suitable TPC track. Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for background and detector effects. The thin slice technique is a new method to measure hadron-argon cross sections possible only due to the combination of the tracking and calorimetry capability of the LArTPC technology. 
Albeit never on argon, the hadronic cross section of pions has been measured before on several different elements in thin target experiments, leading to solid predictions for measurements on argon. Through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the prediction by the Geant4 Bertini Cascade model which are based on data from thin target experiments. On the contrary, cross section measurements for kaons are extremely rare, thus more difficult to model. Our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in disagreement with the Geant4 prediction and provides new experimental data to properly tune the Geant4 models.

@@

The ($\pi^-$-Ar) and the ($K^+$-Ar) total hadronic cross measurements are the first physics results of the LArIAT experiment and will be the basis for the future LArIAT measurements of pion and kaon cross sections in the exclusive channels.

>>>>

This thesis presents two world's first measurements: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1050 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-650 MeV kinetic energy range. The analyses devised for these measurements use both the core elements of LArIAT: beamline and LArTPC. The first step in each analysis is the development of an event selection based on  beamline and TPC information geared towards the identification of the hadron of interest. We then proceed to match the beamline candidate to a suitable TPC track. Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for background contamination and detector effects. The thin slice technique is a new method to measure hadron-argon cross sections possible only due to the combination of the tracking and calorimetry capability of the LArTPC technology. 
Albeit never on argon, the hadronic cross section of pions has been extensively measured before on lighter and heavier different elements in thin target experiments, leading to solid predictions for measurements on argon. Through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the predictions by the Geant4 Bertini Cascade model which are based on data from thin target experiments. On the contrary, cross section measurements for kaons are extremely scarce, thus more difficult to model. Our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in tension with the Geant4 prediction over the explored range of energies and provides new experimental data to properly compare existing interaction models.

@@

The ($\pi^-$-Ar) and the ($K^+$-Ar) total hadronic cross measurements are the first physics results of the LArIAT experiment and will be the basis for the future LArIAT measurements of pion and kaon cross sections in the exclusive channels.
The outcome of these measurements will ultimately enable to quantify and reduce the systematic associated with the hadronic interaction models in neutrino-argon interactions.
